---
title: "Data Analyis_MainAnalysis"
author: "Marshae Nickelberry"
date: "2024-06-24"
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Load packages and datasets
```{r, echo=FALSE}
library(here)
library(knitr)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(gridExtra)
library(survey)
library(tidyr)
library(dplyr)
library(expss)
library(broom)
library(gtsummary)
library(readxl)
library(emmeans)
library(corrplot)

```

## Knit

```{r}
knitr::opts_knit$set(root.dir = here::here())
```

## Functions
```{r}
# Function to exponentiate coefficients and include additional statistics
exp_coeffs <- function(model) {
  # Extract coefficients, standard errors, t-values, and p-values
  coefs <- summary(model)$coefficients
  exp_coefs <- exp(coefs[, 1])  # Exponentiated coefficients
  conf_int <- exp(confint(model))  # Exponentiated confidence intervals
  
  # Combine into a data frame
  exp_results <- data.frame(
    Term = rownames(coefs),
    Estimate = exp_coefs,
    `Std. Error` = coefs[, 2],
    `t value` = coefs[, 3],
    `Pr(>|t|)` = coefs[, 4],
    `2.5 %` = conf_int[, 1],
    `97.5 %` = conf_int[, 2]
  )
  
  return(exp_results)
}

```


#More Data Transformations
```{r, include=FALSE}

FullData_allcyclesFINAL2<-readRDS("FullData_allcyclesFINAL2.rds")

#change units of creatinine (mg/dl)
FullData_allcyclesFINAL2$URXUCR_final <- FullData_allcyclesFINAL2$URXUCR_final * 10000

#dropped those with missing creatinine
FullData_allcyclesFINAL2 = FullData_allcyclesFINAL2[!is.na(FullData_allcyclesFINAL2$URXUCR_final), ] 


table(FullData_allcyclesFINAL2$FOODSEC, useNA="ifany")

#rename variables 
FullData_allcyclesFINAL2$SEX <- factor(FullData_allcyclesFINAL2$SEX, levels = c(1, 2), labels = c("Male", "Female"))



FullData_allcyclesFINAL2$RACEETH <- factor(FullData_allcyclesFINAL2$RACEETH, levels = c(1,2,3,4,5), labels = c("NH White", "NH Black", "Mexican-American", "Other Hispanic", "Other"))

FullData_allcyclesFINAL2$EDUCAT <- factor(FullData_allcyclesFINAL2$EDUCAT, levels=c(1,2,3,4), labels= c("Less than 9th grade", "High school graduate/GED or equivalent equivalent","Some college or AA degree", "College or above")) 

FullData_allcyclesFINAL2$PIRCAT <- factor(FullData_allcyclesFINAL2$PIRCAT, levels=c(1,2,3,4,5), labels= c("Low Poverty","Near Poverty", "Middle Income","Upper-Middle Income", "High Income")) 

FullData_allcyclesFINAL2$FOODSEC <- factor(FullData_allcyclesFINAL2$FOODSEC, levels = c(1, 2, 3, 4), labels= c("Fully Food Secure", "Marginally Food Secure", "Low Food Security", "Very Low Food Secure"))

FullData_allcyclesFINAL2$cycleyr <- factor(FullData_allcyclesFINAL2$cycleyr, levels = c(1,2,3,4,5,6), labels = c("2005-2006","2007-2008","2009-2010","2011-2012","2013-2014", "2015-2016")) 

FullData_allcyclesFINAL2$AGEGROUP <- factor(
  FullData_allcyclesFINAL2$AGEGROUP,
  levels = c("18-31", "32-46", "47-63", "64-85")
)

table(FullData_allcyclesFINAL2$FOODSEC, useNA="ifany")


# FullData_allcyclesFINAL2 <- FullData_allcyclesFINAL2 %>%
#   mutate(PIRCAT = factor(
#     case_when(
#       PIRCAT == "Low Poverty" ~ "Below Poverty",
#       PIRCAT %in% c("Near Poverty", "Middle Income") ~ "Near/Working Class",
#       PIRCAT %in% c("Upper-Middle Income", "High Income") ~ "High Income"
#     ),
#     levels = c("Below Poverty", "Near/Working Class", "High Income")
#   ))



```

log phenols 
```{r}
library(dplyr)



FullData_allcyclesFINAL2 <- FullData_allcyclesFINAL2 %>%
  mutate(
    URXBPH_LOG = log(URXBPH),
    URXBP3_LOG = log(URXBP3),
    URXTRS_LOG = log(URXTRS),
    URXBUP_LOG = log(URXBUP),
    URXEPB_LOG = log(URXEPB),
    URXMPB_LOG = log(URXMPB),
    URXPPB_LOG = log(URXPPB)
  )

saveRDS(FullData_allcyclesFINAL2, "FullData_allcyclesFINAL2.rds")
```


# Survey Package 

now we need to use the survey package to do survey analysis 


```{r}
FullData_allcyclesFINAL2 <- subset(FullData_allcyclesFINAL2, SAMPLE2 == 1)

table(FullData_allcyclesFINAL2$PIRCAT, useNA="ifany")
table(FullData_allcyclesFINAL2$FOODSEC, useNA="ifany")
table(FullData_allcyclesFINAL2$SEX, useNA="ifany")
table(FullData_allcyclesFINAL2$RACEETH, useNA="ifany")
table(FullData_allcyclesFINAL2$AGEGROUP, useNA="ifany")
table(FullData_allcyclesFINAL2$EDUCAT, useNA="ifany")
summary(FullData_allcyclesFINAL2$SURVWEIGHT, useNA="ifany")


saveRDS(FullData_allcyclesFINAL2, "FullData_allcyclesFINAL2.rds")

```


```{r}

FullData_allcyclesFINAL2<-readRDS("FullData_allcyclesFINAL2.rds")

survey <-svydesign(ids = ~SDMVPSU,
          strata = ~SDMVSTRA,
         weights = ~SURVWEIGHT,
         nest = TRUE, 
         data = FullData_allcyclesFINAL2)


```


#TABLE ONE
```{r, echo=FALSE}

#class(survey)
survey %>% 
  tbl_svysummary(
    # Use a character variable here. A factor leads to an error
    by = FOODSEC,
    # Use include to select variables
    include = c(SEX, RIDAGEYR, RACEETH, EDUCAT,
                PIRCAT, cycleyr),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}    ({p}%)"),
    digits = list( all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Variable**",
    all_stat_cols()~"**{level}**")%>%
  modify_caption("Weighted descriptive statistics, by food security status") %>%
  bold_labels()



survey_unweighted<-as.data.frame(survey)
tbl_summary(
  survey_unweighted,
    # Use a character variable here. A factor leads to an error
    by = FOODSEC,
    # Use include to select variables
    include = c(SEX, RIDAGEYR, RACEETH, EDUCAT,
                PIRCAT, cycleyr),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}    ({p}%)"),
    digits = list( all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Variable**",
    all_stat_cols()~"**{level}**")%>%
  modify_caption("Unweighted descriptive statistics, by food security status") %>%
  bold_labels()%>%
  add_overall()
```
```{r}
table(FullData_allcyclesFINAL2$FOODSEC, useNA="ifany")
```


# GEOMETRIC MEANS

Log transformed values instead for geometric mean reporting.

Realized that NHANES may have updated check log for details 

```{r}

log_means_set2<- svymean(~URXBPH_LOG+URXBP3_LOG+URXTRS_LOG+URXBUP_LOG+URXEPB_LOG+URXMPB_LOG+URXPPB_LOG, design=survey, na.rm = TRUE)
print(log_means_set2) 

#exponetiate


log_means<- coef(log_means_set2) # log-scale means
geomeans<-exp(log_means)
log_se <- SE(log_means_set2)  # standard errors on log scale


geomean_table <- data.frame(
  variable = names(log_means),
  geomean = geomeans
)


print(geomean_table)


```


#TABLE TWO
```{r, echo=FALSE}
# Step 1 – Run svyquantile for each phenol and save


svyquantile(~URXBPH_LOG, design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXBP3_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXTRS_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXBUP_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXEPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXMPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXPPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)



```



```{r}
table(FullData_allcyclesFINAL2$URDBPHLC, useNA = "ifany") #9287/(9287+807)
table(FullData_allcyclesFINAL2$URDBP3LC, useNA = "ifany") #9706/(9706+388)
table(FullData_allcyclesFINAL2$URDTRSLC, useNA = "ifany") #7546/(7546+2548)
table(FullData_allcyclesFINAL2$URDBUPLC, useNA = "ifany")#3716/(3716+6378)
table(FullData_allcyclesFINAL2$URDEPBLC,useNA = "ifany")#4999/(4999+5095)
table(FullData_allcyclesFINAL2$URDMPBLC, useNA = "ifany" ) #10010/(10010+84)
table(FullData_allcyclesFINAL2$URDPPBLC, useNA = "ifany" ) #9607/(9607+487)





```

```{r}
sum(!is.na(FullData_allcyclesFINAL2$URDBPHLC))
sum(!is.na(FullData_allcyclesFINAL2$URDBP3LC))
sum(!is.na(FullData_allcyclesFINAL2$URDTRSLC))
sum(!is.na(FullData_allcyclesFINAL2$URDBUPLC))
sum(!is.na(FullData_allcyclesFINAL2$URDEPBLC))
sum(!is.na(FullData_allcyclesFINAL2$URDMPBLC))
sum(!is.na(FullData_allcyclesFINAL2$URDPPBLC))
```

#CORRELATION PLOT 
```{r}

FullData_allcyclesFINAL2<-readRDS("FullData_allcyclesFINAL2.rds")

names(FullData_allcyclesFINAL2)
#LOG
vars_LOG<-FullData_allcyclesFINAL2%>% 
  select(c(URXBPH_LOG, URXBP3_LOG, URXTRS_LOG, URXEPB_LOG, URXMPB_LOG, URXPPB_LOG))

new_names <- c("Bisphenol A", "Benzophenone-3", "Triclosan", "Ethyl Paraben", "Methyl Paraben", "Propyl Paraben")

cor_mat_LOG<-cor(vars_LOG, use = "pairwise.complete.obs", method="spearman")

colnames(cor_mat_LOG) <- new_names
rownames(cor_mat_LOG) <- new_names

corrplot(cor_mat_LOG, type="full", order= "original", col = brewer.pal(n=8, name= "RdYlBu"), tl.srt = 45, addCoef.col = "black", number.cex = 0.7)

```






# MAIN ANALYSIS

## Unadjusted analysis

```{r, echo=FALSE}
#URXBPH_LOG

mod1_unadj<-svyglm(URXBPH_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXBP3_LOG
mod2_unadj<-svyglm(URXBP3_LOG~(FOODSEC) + URXUCR_final , design=survey)

#URXTRS_LOG
mod3_unadj<-svyglm(URXTRS_LOG~(FOODSEC) + URXUCR_final , design=survey)

#URXBUP_LOG
mod4_unadj<-svyglm(URXBUP_LOG~(FOODSEC) + URXUCR_final,  design=survey)

#URXEPB_LOG

mod5_unadj<-svyglm(URXEPB_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXMPB_LOG
mod6_unadj<-svyglm(URXMPB_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXPPB_LOG
mod7_unadj<-svyglm(URXPPB_LOG~(FOODSEC) + URXUCR_final, design=survey)



# Apply the exp function to each model and print the results
models_unadj<- list(mod1_unadj, mod2_unadj, mod3_unadj, mod4_unadj, mod5_unadj, 
               mod6_unadj, mod7_unadj )

names(models_unadj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG",
                    "URXMPB_LOG", "URXPPB_LOG")

exp_unadj<- lapply(models_unadj, exp_coeffs)

# Combine all dataframes into one
unadj_combined_df <- bind_rows(exp_unadj)

#write csv 

write.csv(unadj_combined_df, "Results/Unadjusted_Models.csv")
```

## Adjusted (age, sex, education, survey year, race, FIP)

```{r, echo=FALSE}
#URXBPH_LOG
mod1c_adj<- svyglm(URXBPH_LOG ~ (FOODSEC) + URXUCR_final + (SEX) + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr),design = survey)



#URXBP3_LOG

mod2c_adj<- svyglm(URXBP3_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXTRS_LOG

mod3c_adj<- svyglm(URXTRS_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#MCPP
mod4c_adj<- svyglm(URXBUP_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXEPB_LOG

mod5c_adj<- svyglm(URXEPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXMPB_LOG
mod6c_adj<- svyglm(URXMPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (cycleyr), design = survey)



#URXPPB_LOG
mod7c_adj<- svyglm(URXPPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



# Apply the exp function to each model and print the results
models_adj <- list(mod1c_adj, mod2c_adj, mod3c_adj, mod4c_adj, mod5c_adj, mod6c_adj, mod7c_adj)

names(models_adj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG", "URXMPB_LOG", "URXPPB_LOG" )

exp_adj<- lapply(models_adj, exp_coeffs)

# Combine all dataframes into one
adj_combined_dfs<- bind_rows(exp_adj)

#write csv 
write.csv(adj_combined_dfs, "Results/Adjusted_Models2.csv")
```


with age as continuous just in case
```{r, echo=FALSE}

#URXBPH_LOG
mod1c_adj<- svyglm(URXBPH_LOG ~ (FOODSEC) + URXUCR_final + (SEX) + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr),design = survey)



#URXBP3_LOG

mod2c_adj<- svyglm(URXBP3_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXTRS_LOG

mod3c_adj<- svyglm(URXTRS_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#MCPP
mod4c_adj<- svyglm(URXBUP_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXEPB_LOG

mod5c_adj<- svyglm(URXEPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



#URXMPB_LOG
mod6c_adj<- svyglm(URXMPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (cycleyr), design = survey)




#URXPPB_LOG
mod7c_adj<- svyglm(URXPPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIRCAT) + (cycleyr), design = survey)



# Apply the exp function to each model and print the results
models_adj <- list(mod1c_adj, mod2c_adj, mod3c_adj, mod4c_adj, mod5c_adj, mod6c_adj, mod7c_adj)

names(models_adj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG", "URXMPB_LOG", "URXPPB_LOG" )

exp_adj<- lapply(models_adj, exp_coeffs)

# Combine all dataframes into one
adj_combined_dfs<- bind_rows(exp_adj)

#write csv 
write.csv(adj_combined_dfs, "Results/Adjusted_Models2b.csv")
```

##Interaction 

```{r}

FullData_allcyclesFINAL2$food_security <- as.numeric(FullData_allcyclesFINAL2$FOODSEC)
class(FullData_allcyclesFINAL2$food_security)


saveRDS(FullData_allcyclesFINAL2, "FullData_allcyclesFINAL2.rds")
```

```{r, include=FALSE}
survey <-svydesign(ids = ~SDMVPSU,
          strata = ~SDMVSTRA,
         weights = ~SURVWEIGHT,
         nest = TRUE, 
         data = FullData_allcyclesFINAL2)

summary(survey)

summary(FullData_allcyclesFINAL2$SURVWEIGHT, useNA="ifany")


```

###Sex 
```{r}

# URXBPH_LOG p= 0.016661
I.genmod1 <- svyglm(URXBPH_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod1, ~food_security:SEX)

# URXBP3_LOG p= 0.34129 
I.genmod2 <- svyglm(URXBP3_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod2, ~food_security:SEX)

# URXTRS_LOG p= 0.88496 
I.genmod3 <- svyglm(URXTRS_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod3, ~food_security:SEX)

# URXBUP_LOG p= 6.273e-09 
I.genmod4 <- svyglm(URXBUP_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod4, ~food_security:SEX)

# URXEPB_LOG p=0.0036593
I.genmod5 <- svyglm(URXEPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod5, ~food_security:SEX)

# URXMPB_LOG p=0.0013114
I.genmod6 <- svyglm(URXMPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod6, ~food_security:SEX)

# URXPPB_LOG p=0.0035299
I.genmod7 <- svyglm(URXPPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.genmod7, ~food_security:SEX)

```

#####Strata estimates 
```{r}
# URXBPH_LOG
I.genBmod1 <- svyglm(URXBPH_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXBP3_LOG
I.genBmod2 <- svyglm(URXBP3_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXTRS_LOG
I.genBmod3 <- svyglm(URXTRS_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXBUP_LOG
I.genBmod4 <- svyglm(URXBUP_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXEPB_LOG
I.genBmod5 <- svyglm(URXEPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXMPB_LOG
I.genBmod6 <- svyglm(URXMPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)

# URXPPB_LOG
I.genBmod7 <- svyglm(URXPPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIRCAT + cycleyr,
                     design = survey)
```


```{r}
emmeans::emm_options(rg.limit = 30000)

emms <- emmeans(I.genBmod1, ~ FOODSEC | SEX)
contrast(emms, method = "revpairwise")
```


```{r}

GENmodels_list<-list(I.genBmod1, I.genBmod2, I.genBmod3, I.genBmod4, I.genBmod5, I.genBmod6, I.genBmod7)
GENmod_names<-c("BPA", "BP3", "TRS", "BUP", "EPB", "MPB", "PPB")

emmeans::emm_options(rg.limit = 30000)

results <- lapply(seq_along(GENmodels_list), function(i) {
  model <- GENmodels_list[[i]]
  mod_name <- GENmod_names[i]
  
  emms <- emmeans(model, ~ FOODSEC | SEX)
  cont <- contrast(emms, method = "revpairwise")
  cont_summary <- summary(cont, infer = TRUE)
  
  # Exponentiate contrasts
  cont_summary$exp.estimate     <- exp(cont_summary$estimate)
  cont_summary$exp.lower.CL     <- exp(cont_summary$lower.CL)
  cont_summary$exp.upper.CL     <- exp(cont_summary$upper.CL)
  
  cont_summary$model <- mod_name
  cont_summary
})

GENall_results <- do.call(rbind, results)
write.csv(GENall_results, "Results/GENstrata_results.csv", row.names=FALSE)
```


###Race/Ethnicity 


```{r}
# BPA_LOG  p= 0.10837  
I.REmod1 <- svyglm(URXBPH_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey )
regTermTest(I.REmod1, ~food_security:RACEETH) 

# BP3_LOG  p= 0.057371
I.REmod2 <- svyglm(URXBP3_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.REmod2, ~food_security:RACEETH)

# TRS_LOG p= 0.081281
I.REmod3 <- svyglm(URXTRS_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.REmod3, ~food_security:RACEETH) 

# BUP_LOG  p= 0.002877
I.REmod4 <- svyglm(URXBUP_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)
regTermTest(I.REmod4, ~food_security:RACEETH) 

# EPB_LOG p= 0.041265
I.REmod5 <- svyglm(URXEPB_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

regTermTest(I.REmod5, ~food_security:RACEETH)

# MPB_LOG p= 0.0059458
I.REmod6 <- svyglm(URXMPB_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

regTermTest(I.REmod6, ~food_security:RACEETH)

# PPB_LOG p= 0.0041702
I.REmod7 <- svyglm(URXPPB_LOG ~ food_security * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

regTermTest(I.REmod7, ~food_security:RACEETH)

```


#####Strata estimates 
```{r}
# BPA_LOG  
I.REBmod1 <- svyglm(URXBPH_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey )

# BP3_LOG 
I.REBmod2 <- svyglm(URXBP3_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

# TRS_LOG 
I.REBmod3 <- svyglm(URXTRS_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

# BUP_LOG  
I.REBmod4 <- svyglm(URXBUP_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)

# EPB_LOG
I.REBmod5 <- svyglm(URXEPB_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)


# MPB_LOG
I.REBmod6 <- svyglm(URXMPB_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)


# PPB_LOG p= 0.0041702
I.REBmod7 <- svyglm(URXPPB_LOG ~ FOODSEC * RACEETH + URXUCR_final + AGEGROUP + EDUCAT + SEX + PIRCAT + cycleyr,
                    design = survey)



```


```{r}

REmodels_list<-list(I.REBmod1, I.REBmod2, I.REBmod3, I.REBmod4, I.REBmod5, I.REBmod6, I.REBmod7)
REmod_names<-c("BPA", "BP3", "TRS", "BUP", "EPB", "MPB", "PPB")

emmeans::emm_options(rg.limit = 30000)

results <- lapply(seq_along(REmodels_list), function(i) {
  model <- REmodels_list[[i]]
  mod_name <- REmod_names[i]
  
  emms <- emmeans(model, ~ FOODSEC | RACEETH)
  cont <- contrast(emms, method = "revpairwise")
  cont_summary <- summary(cont, infer = TRUE)
  
  # Exponentiate contrasts
  cont_summary$exp.estimate     <- exp(cont_summary$estimate)
  cont_summary$exp.lower.CL     <- exp(cont_summary$lower.CL)
  cont_summary$exp.upper.CL     <- exp(cont_summary$upper.CL)
  
  cont_summary$model <- mod_name
  cont_summary
})

REall_results <- do.call(rbind, results)
write.csv(REall_results, "Results/REstrata_results.csv", row.names=FALSE)
```


#PLOTS
###Model 3 adjusted signifcant findings
```{r}

RM_GGPlots <-read.csv("Results/RM_GGPlots.csv")
names(RM_GGPlots)
names(RM_GGPlots) <- c("Model", "FOODSEC", "Food Security", "% Difference", "% Lower", "% Higher",  "P_value")
brewer_pal <- brewer.pal(3, "Set1")
range(RM_GGPlots$`% Lower`, na.rm=TRUE)
range(RM_GGPlots$`% Higher`, na.rm=TRUE)



# Helper theme chunk to reuse in each plot
custom_theme <-
  theme(
    plot.title = element_text(hjust = 0.5, face="bold", size=18),
    panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),     
    plot.margin = margin(t = 10, r = 10, b = 25, l = 10)  # 
  )

# BPA
BPA_adj3 <- RM_GGPlots %>%
  filter(Model == "BPA", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
  ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.3, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BPA",
    x = "Food Security Levels",
    y = "% Difference in Concentrations") +
  labs(title = "BPA", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)  + 
  scale_y_continuous(limits = c(-42, 25))


# BP3
BP3_adj3 <- RM_GGPlots %>%
  filter(Model == "BP3", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
      ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "BP3", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+   scale_y_continuous(limits = c(-42, 25))


# TRS
TRS_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "TRS", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "TRS", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+   
  scale_y_continuous(limits = c(-42, 25))


# BUP
BUP_adj3 <- RM_GGPlots %>%
  filter(Model == "BUP", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
  ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "BUP", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
  scale_y_continuous(limits = c(-42, 25))


# EPB
EPB_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "EPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "EPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))


# MPB
MPB_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "MPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "MPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))


#PPB
PPB_adj3 <- RM_GGPlots %>%
  filter(Model == "PPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
 ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "PPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))

```


```{r}

figure1<- ggarrange(
  BPA_adj3 + theme(legend.position="none"), 
  BP3_adj3 + theme(legend.position="none"), 
  TRS_adj3 + theme(legend.position="none"),
  #BUP_adj3 + theme(legend.position="none"), 
  EPB_adj3 + theme(legend.position="none"), 
  MPB_adj3 + theme(legend.position="none"), 
  PPB_adj3 , # Leave legend ON only here
  ncol = 3, nrow = 3,
  labels = c("A", "B", "C", "D", "E", "F"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure1

```


## Stratified by gender plots

```{r}
GENstrata<-read.csv("Results/GENstrata2.csv")
brewer_pal <- brewer.pal(3, "Set1")

range(GENstrata$`X.Lower`, na.rm=TRUE) #-65 
range(GENstrata$`X.Higher`, na.rm=TRUE) #55



#create vector with race/ethnicity labels
sex_labels<- c("Male", "Female")
sex_colors <- c("Male" = brewer_pal[1], "Female" = brewer_pal[2])
names(sex_colors)<-sex_labels
sex_shapes <- c("Male" = 16, "Female" = 17) # Circle, triangle
names(sex_shapes)<-sex_labels



# BPA
BPA_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "BPA", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BPA",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BPA_adjEMM

# BP3
BP3_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "BP3", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BP3",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BP3_adjEMM

# TRS
TRS_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "TRS", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "TRS",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))  +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

TRS_adjEMM

# BUP
BUP_adjEMM<- ggplot(
  data = GENstrata %>%
    filter(model == "BUP", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BUP",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BUP_adjEMM

# EPB

EPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "EPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "EPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

EPB_adjEMM

# MPB

MPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "MPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "MPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))  +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

# PPB
PPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "PPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "PPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                  labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )
PPB_adjEMM


```


```{r}



# BUP_adjEMM <- BUP_adjEMM +
#   annotate("text",
#            x = 2.5, y = 55,
#            label = bquote(P[interaction] == .(formatC(0.0001, digits=4, format="f"))),
#            size=5, hjust=0.5)


p_int_BPA  <- 0.02
p_int_BP3  <- 0.34
p_int_TRS  <- 0.88
p_int_EPB  <- 0.004
p_int_MPB  <- 0.001
p_int_PPB  <- 0.003

BPA_adjEMM <- BPA_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.2f", p_int_BPA),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )

BP3_adjEMM <- BP3_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.2f", p_int_BP3),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )

TRS_adjEMM <- TRS_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.2f", p_int_TRS),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )

EPB_adjEMM <- EPB_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.3f", p_int_EPB),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )

MPB_adjEMM <- MPB_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.3f", p_int_MPB),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )

PPB_adjEMM <- PPB_adjEMM +
  annotate(
    "text",
    x = 2.5, y = 55,
    label = sprintf("P[interaction]==%.3f", p_int_PPB),
    parse  = TRUE,
    size = 5, hjust = 0.5
  )


PPB_adjEMM

figure2<- ggarrange(
  BPA_adjEMM + theme(legend.position="none"),
  BP3_adjEMM + theme(legend.position="none"), 
  TRS_adjEMM + theme(legend.position="none"), 
  EPB_adjEMM + theme(legend.position="none"), 
  MPB_adjEMM + theme(legend.position="none"), 
  PPB_adjEMM,  # Leave legend ON only here
  ncol = 3, nrow = 3,
  labels = c("A", "B", "C", "D", "E", "F"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure2

ggexport(figure2, filename = "Results/figure2.pdf")
```


##Race Stratified
```{r}
RACEstrata<-read.csv("Results/REstrata2.csv")
#factor first to make sure correct order is presented 
RACEstrata$RACEETH <- factor(
  RACEstrata$RACEETH,
  levels = c("NH White", "NH Black", "Mexican-American", "Other Hispanic", "Other")
)

#create vector with race/ethnicity labels
raceeth_labels<- c("NH White", "NH Black", "Mexican-American", "Other Hispanic", "Other")
raceeth_colors<- brewer.pal(n = length(raceeth_labels), name = "Set1")
names(raceeth_colors)<-raceeth_labels
raceeth_shapes<- c(16, 17, 15, 18, 8) # Circle, triangle, square, diamond, star
names(raceeth_shapes)<-raceeth_labels


range(RACEstrata$`X.Higher`, na.rm=TRUE) # 251.52; 255
range(RACEstrata$`X.Lower`, na.rm=TRUE) #-83.14; 85



# MnBP
BPA_adjREMM <- ggplot(
  data = RACEstrata %>%
    filter(model == "BPA", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  # adds points (scatterplot dots) to ggplot mapping data from what was put into aes (for x, y, and combinations of color and group)
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  # position dodge dictates how far/spread (width). The size and transparency of the points: 0 is fully transparent and 1 is fully opaque

  # adds my error bars and also has position dodge options but in this case (width is for spread) and width is for the cross bar width at the top of each error bar
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
                
  # adds a reference line, at the y intercept, makes it zero and assigns color
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  

  # labs sets labels 
  labs(
    title = "BPA",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +

  # controls what colors are assigned and manual allows you to choose which colors. 
  # You made a vector for everyone already and breaks specifies order of groups in legends and ensures only those labels appear
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  # similar as above 
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +

  # ensures all levels are shown even if some are missing,
  # assigns labels
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "1(Ref)", "2"="2", "3"="3", "4"="4")) +
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), # hjust centers the title, makes it bold and increases the font size
     panel.grid.major.y = element_line(color = "grey90"),           # uncomment if you want faint horizontal gridlines
    panel.grid.minor = element_blank(),                            # uncomment if you want no minor gridlines
    axis.line = element_line(color = "black"),                       # draws black lines along the axes; helps frame the plot 
    legend.position = "bottom",                                      # places legend at bottom
    legend.title = element_text(size = 15, face = "bold"),           # legend title 
    legend.text = element_text(size = 14),                           # legend text size
    legend.box = "horizontal",                                       # put legend horizontally beneath facets
    legend.key.size = unit(1.5, "lines"),                            # focuses on icons and their sizes
    strip.text = element_text(face = "bold", size = 16)              # makes facet title text the header for each small multiple
  ) +

  # legend title above, all items in one row
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )
BPA_adjREMM

# BP3

BP3_adjREMM <- ggplot(
  data = RACEstrata %>%
    filter(model == "BP3", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(
    title = "BP3",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))+
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )



# TRS
TRS_adjREMM<- ggplot(
  data = RACEstrata %>%
    filter(model == "TRS", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(
    title = "TRS",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))+
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )
TRS_adjREMM


# BUP
BUP_adjREMM<-ggplot(
  data = RACEstrata %>%
    filter(model == "BUP", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(

    title = "BUP",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                  labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))+
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )
BUP_adjREMM



# EPB
EPB_adjREMM<-ggplot(
  data = RACEstrata %>%
    filter(model == "EPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(
    title = "EPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )
EPB_adjREMM

# MPB
MPB_adjREMM<-ggplot(
  data = RACEstrata %>%
    filter(model == "MPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(
    title = "MPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )

# PPB
PPB_adjREMM<-ggplot(
  data = RACEstrata %>%
    filter(model == "PPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(
    x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
    y = X.Difference,
    color = RACEETH,
    shape = RACEETH
  )
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(
    title = "PPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Race/Ethnicity", # title for color legend 
    shape = "Race/Ethnicity" # title for shape legend
  ) +
  scale_color_manual(values = raceeth_colors, breaks = raceeth_labels) +
  scale_shape_manual(values = raceeth_shapes, breaks = raceeth_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-85, 255))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)
  )






```




```{r}

# BUP_adjREMM <- BUP_adjREMM +
#   annotate("text",
#            x = 2.5, y = 250,
#           label = bquote(P[interaction] == .(formatC(0.003, digits=3, format="f"))),
#            size=4, hjust=0.5)

# interaction p-values
p_int_BPA_REMM <- 0.11
p_int_BP3_REMM <- 0.06
p_int_TRS_REMM <- 0.08
p_int_BUP_REMM <- 0.003
p_int_EPB_REMM <- 0.04
p_int_MPB_REMM <- 0.006
p_int_PPB_REMM <- 0.004



BPA_adjREMM <- BPA_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.2f", p_int_BPA_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

BP3_adjREMM <- BP3_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.2f", p_int_BP3_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

TRS_adjREMM <- TRS_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.2f", p_int_TRS_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

BUP_adjREMM <- BUP_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.3f", p_int_BUP_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

EPB_adjREMM <- EPB_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.2f", p_int_EPB_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

MPB_adjREMM <- MPB_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.3f", p_int_MPB_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

PPB_adjREMM <- PPB_adjREMM +
  annotate(
    "text",
    x = 2.5, y = 250,
    label = sprintf("P[interaction]==%.3f", p_int_PPB_REMM),
    parse  = TRUE,
    size   = 4, hjust = 0.5
  )

figure3a<- ggarrange(
  BPA_adjREMM + theme(legend.position="none"), 
  BP3_adjREMM+ theme(legend.position="none"), 
  TRS_adjREMM,
# Leave legend ON only here
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure3a

figure3b<- ggarrange(
  #BUP_adjREMM + theme(legend.position="none"), 
  EPB_adjREMM + theme(legend.position="none"), 
  MPB_adjREMM + theme(legend.position="none"),
  PPB_adjREMM, 
# Leave legend ON only here
  ncol = 2, nrow = 2,
  labels = c("A", "B", "C"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure3b

ggexport(figure3a, filename = "Results/figure3a.pdf")

ggexport(figure3b, filename = "Results/figure3b.pdf")
```





