---
title: "Data Analyis_MainAnalysis"
author: "Marshae Nickelberry"
date: "2024-06-24"
output:
  html_document: default
  pdf_document: default
editor_options: 
  markdown: 
    wrap: 72
---

# Load packages and datasets
```{r, echo=FALSE}
library(here)
library(knitr)
library(ggplot2)
library(ggpubr)
library(RColorBrewer)
library(gridExtra)
library(survey)
library(tidyr)
library(dplyr)
library(expss)
library(broom)
library(plyr)
library(gtsummary)
library(readxl)
library(emmeans)
library(corrplot)

```

## Knit

```{r}
knitr::opts_knit$set(root.dir = here::here())
```

## Functions
```{r}
# Function to exponentiate coefficients and include additional statistics
exp_coeffs <- function(model) {
  # Extract coefficients, standard errors, t-values, and p-values
  coefs <- summary(model)$coefficients
  exp_coefs <- exp(coefs[, 1])  # Exponentiated coefficients
  conf_int <- exp(confint(model))  # Exponentiated confidence intervals
  
  # Combine into a data frame
  exp_results <- data.frame(
    Term = rownames(coefs),
    Estimate = exp_coefs,
    `Std. Error` = coefs[, 2],
    `t value` = coefs[, 3],
    `Pr(>|t|)` = coefs[, 4],
    `2.5 %` = conf_int[, 1],
    `97.5 %` = conf_int[, 2]
  )
  
  return(exp_results)
}

```


#More Data Transformations
```{r, include=FALSE}
FullData_allcyclesFINAL2<-read.csv("FullData_allcyclesFINAL2.csv")

#FullData_allcyclesFINAL2<-readRDS("FullData_allcyclesFINAL2.rds")

#change units of creatinine (mg/dl)
FullData_allcyclesFINAL2$URXUCR_final <- FullData_allcyclesFINAL2$URXUCR_final * 10000

#dropped those with missing creatinine
FullData_allcyclesFINAL2 = FullData_allcyclesFINAL2[!is.na(FullData_allcyclesFINAL2$URXUCR_final), ] 


table(FullData_allcyclesFINAL2$FOODSEC, useNA="ifany")

#rename variables 
FullData_allcyclesFINAL2$SEX <- factor(FullData_allcyclesFINAL2$SEX, levels = c(1, 2), labels = c("Male", "Female"))



FullData_allcyclesFINAL2$RACEETH <- factor(FullData_allcyclesFINAL2$RACEETH, levels = c(1,2,3,4,5), labels = c("NH White", "NH Black", "Mexican-American", "Other Hispanic", "Other"))

FullData_allcyclesFINAL2$EDUCAT <- factor(FullData_allcyclesFINAL2$EDUCAT, levels=c(1,2,3,4), labels= c("Less than 9th grade", "High school graduate/GED or equivalent equivalent","Some college or AA degree", "College or above")) 

FullData_allcyclesFINAL2$PIRCAT <- factor(FullData_allcyclesFINAL2$PIRCAT, levels=c(1,2,3,4,5), labels= c("Low Poverty","Near Poverty", "Middle Income","Upper-Middle Income", "High Income")) 

FullData_allcyclesFINAL2$FOODSEC <- factor(FullData_allcyclesFINAL2$FOODSEC, levels = c(1, 2, 3, 4), labels= c("Fully Food Secure", "Marginally Food Secure", "Low Food Security", "Very Low Food Secure"))

FullData_allcyclesFINAL2$cycleyr <- factor(FullData_allcyclesFINAL2$cycleyr, levels = c(1,2,3,4,5,6), labels = c("2005-2006","2007-2008","2009-2010","2011-2012","2013-2014", "2015-2016")) 

FullData_allcyclesFINAL2$AGEGROUP <- factor(
  FullData_allcyclesFINAL2$AGEGROUP,
  levels = c("18-31", "32-46", "47-63", "64-85")
)

table(FullData_allcyclesFINAL2$FOODSEC)


FullData_allcyclesFINAL2 <- FullData_allcyclesFINAL2 %>%
  mutate(PIR_POLICY = factor(
    case_when(
      PIRCAT == "Low Poverty" ~ "Below Poverty",
      PIRCAT %in% c("Near Poverty", "Middle Income") ~ "Near/Working Class",
      PIRCAT %in% c("Upper-Middle Income", "High Income") ~ "High Income"
    ),
    levels = c("Below Poverty", "Near/Working Class", "High Income")
  ))


```

log phenols 
```{r}
library(dplyr)

FullData_allcyclesFINAL2 <- FullData_allcyclesFINAL2 %>%
  mutate(
    URXBPH_LOG = log(URXBPH),
    URXBP3_LOG = log(URXBP3),
    URXTRS_LOG = log(URXTRS),
    URXBUP_LOG = log(URXBUP),
    URXEPB_LOG = log(URXEPB),
    URXMPB_LOG = log(URXMPB),
    URXPPB_LOG = log(URXPPB)
  )
```


# Survey Package 

now we need to use the survey package to do survey analysis 


```{r}
FullData_allcyclesFINAL2 <- subset(FullData_allcyclesFINAL2, SAMPLE2 == 1)


```


```{r}
survey <-svydesign(ids = ~SDMVPSU,
          strata = ~SDMVSTRA,
         weights = ~SURVWEIGHT,
         nest = TRUE, 
         data = FullData_allcyclesFINAL2)


```



```{r, echo=FALSE}

#class(survey)
survey %>% 
  tbl_svysummary(
    # Use a character variable here. A factor leads to an error
    by = FOODSEC,
    # Use include to select variables
    include = c(SEX, AGEGROUP, RACEETH, EDUCAT,
                PIR_POLICY, cycleyr),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}    ({p}%)"),
    digits = list( all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Variable**",
    all_stat_cols()~"**{level}**")%>%
  modify_caption("Weighted descriptive statistics, by food security status") %>%
  bold_labels()



survey_unweighted<-as.data.frame(survey)
tbl_summary(
  survey_unweighted,
    # Use a character variable here. A factor leads to an error
    by = FOODSEC,
    # Use include to select variables
    include = c(SEX, AGEGROUP, RACEETH, EDUCAT,
                PIR_POLICY, cycleyr),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",
                     all_categorical() ~ "{n}    ({p}%)"),
    digits = list( all_categorical() ~ c(0, 1))
  ) %>%
  modify_header(
    label = "**Variable**",
    all_stat_cols()~"**{level}**")%>%
  modify_caption("Weighted descriptive statistics, by food security status") %>%
  bold_labels()%>%
  add_overall()
```


# GEOMETRIC MEANS

Log transformed values instead for geometric mean reporting.

Realized that NHANES may have updated check log for details 

```{r}

log_means_set2<- svymean(~URXBPH_LOG+URXBP3_LOG+URXTRS_LOG+URXBUP_LOG+URXEPB_LOG+URXMPB_LOG+URXPPB_LOG, design=survey, na.rm = TRUE)
print(log_means_set2) 

#exponetiate


log_means<- coef(log_means_set2) # log-scale means
geomeans<-exp(log_means)
log_se <- SE(log_means_set2)  # standard errors on log scale


geomean_table <- data.frame(
  variable = names(log_means),
  geomean = geomeans
)


print(geomean_table)


```


#TABLE TWO
```{r, echo=FALSE}
# Step 1 – Run svyquantile for each phenol and save


svyquantile(~URXBPH_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXBP3_LOG, design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXTRS_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXBUP_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXEPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)

svyquantile(~URXMPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)
svyquantile(~URXPPB_LOG,  design = survey, quantiles = c(0.25, 0.75), na.rm = TRUE)



```



# MAIN ANALYSIS

## Unadjusted analysis

```{r, echo=FALSE}
#URXBPH_LOG

mod1_unadj<-svyglm(URXBPH_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXBP3_LOG
mod2_unadj<-svyglm(URXBP3_LOG~(FOODSEC) + URXUCR_final , design=survey)

#URXTRS_LOG
mod3_unadj<-svyglm(URXTRS_LOG~(FOODSEC) + URXUCR_final , design=survey)

#URXBUP_LOG
mod4_unadj<-svyglm(URXBUP_LOG~(FOODSEC) + URXUCR_final,  design=survey)

#URXEPB_LOG

mod5_unadj<-svyglm(URXEPB_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXMPB_LOG
mod6_unadj<-svyglm(URXMPB_LOG~(FOODSEC) + URXUCR_final, design=survey)

#URXPPB_LOG
mod7_unadj<-svyglm(URXPPB_LOG~(FOODSEC) + URXUCR_final, design=survey)



# Apply the exp function to each model and print the results
models_unadj<- list(mod1_unadj, mod2_unadj, mod3_unadj, mod4_unadj, mod5_unadj, 
               mod6_unadj, mod7_unadj )

names(models_unadj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG",
                    "URXMPB_LOG", "URXPPB_LOG")

exp_unadj<- lapply(models_unadj, exp_coeffs)

# Combine all dataframes into one
unadj_combined_df <- bind_rows(exp_unadj)

#write csv 

write.csv(unadj_combined_df, "Results/Unadjusted_Models3.csv")
```

## Adjusted (age, sex, education, survey year, race, FIP)

```{r, echo=FALSE}
#URXBPH_LOG
mod1c_adj<- svyglm(URXBPH_LOG ~ (FOODSEC) + URXUCR_final + (SEX) + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr),design = survey)



#URXBP3_LOG

mod2c_adj<- svyglm(URXBP3_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXTRS_LOG

mod3c_adj<- svyglm(URXTRS_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#MCPP
mod4c_adj<- svyglm(URXBUP_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXEPB_LOG

mod5c_adj<- svyglm(URXEPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXMPB_LOG
mod6c_adj<- svyglm(URXMPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (cycleyr), design = survey)



#URXPPB_LOG
mod7c_adj<- svyglm(URXPPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (AGEGROUP) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



# Apply the exp function to each model and print the results
models_adj <- list(mod1c_adj, mod2c_adj, mod3c_adj, mod4c_adj, mod5c_adj, mod6c_adj, mod7c_adj)

names(models_adj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG", "URXMPB_LOG", "URXPPB_LOG" )

exp_adj<- lapply(models_adj, exp_coeffs)

# Combine all dataframes into one
adj_combined_dfs<- bind_rows(exp_adj)

#write csv 
write.csv(adj_combined_dfs, "Results/Adjusted_Models2.csv")
```



```{r, echo=FALSE}

#URXBPH_LOG
mod1c_adj<- svyglm(URXBPH_LOG ~ (FOODSEC) + URXUCR_final + (SEX) + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr),design = survey)



#URXBP3_LOG

mod2c_adj<- svyglm(URXBP3_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXTRS_LOG

mod3c_adj<- svyglm(URXTRS_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#MCPP
mod4c_adj<- svyglm(URXBUP_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXEPB_LOG

mod5c_adj<- svyglm(URXEPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



#URXMPB_LOG
mod6c_adj<- svyglm(URXMPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (cycleyr), design = survey)



#URXPPB_LOG
mod7c_adj<- svyglm(URXPPB_LOG ~ (FOODSEC) + URXUCR_final + SEX + (RIDAGEYR) +  (EDUCAT)+ (RACEETH) + (PIR_POLICY) + (cycleyr), design = survey)



# Apply the exp function to each model and print the results
models_adj <- list(mod1c_adj, mod2c_adj, mod3c_adj, mod4c_adj, mod5c_adj, mod6c_adj, mod7c_adj)

names(models_adj) <- c("URXBPH_LOG", "URXBP3_LOG", "URXTRS_LOG", "URXBUP_LOG", "URXEPB_LOG", "URXMPB_LOG", "URXPPB_LOG" )

exp_adj<- lapply(models_adj, exp_coeffs)

# Combine all dataframes into one
adj_combined_dfs<- bind_rows(exp_adj)

#write csv 
write.csv(adj_combined_dfs, "Results/Adjusted_Models2b.csv")
```

##Interaction 

```{r}

FullData_allcyclesFINAL2$food_security <- as.numeric(FullData_allcyclesFINAL2$FOODSEC)
class(FullData_allcyclesFINAL2$food_security)


saveRDS(FullData_allcyclesFINAL2, "FullData_allcyclesFINAL2.rds")
```

```{r, include=FALSE}
survey <-svydesign(ids = ~SDMVPSU,
          strata = ~SDMVSTRA,
         weights = ~SURVWEIGHT,
         nest = TRUE, 
         data = FullData_allcyclesFINAL2)

summary(survey)



```

###Sex 
```{r}

# URXBPH_LOG
I.genmod1 <- svyglm(URXBPH_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod1, ~food_security:SEX)

# URXBP3_LOG
I.genmod2 <- svyglm(URXBP3_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod2, ~food_security:SEX)

# URXTRS_LOG
I.genmod3 <- svyglm(URXTRS_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod3, ~food_security:SEX)

# URXBUP_LOG
I.genmod4 <- svyglm(URXBUP_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod4, ~food_security:SEX)

# URXEPB_LOG
I.genmod5 <- svyglm(URXEPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod5, ~food_security:SEX)

# URXMPB_LOG
I.genmod6 <- svyglm(URXMPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod6, ~food_security:SEX)

# URXPPB_LOG
I.genmod7 <- svyglm(URXPPB_LOG ~ food_security * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                    design = survey)
regTermTest(I.genmod7, ~food_security:SEX)

```

#####Strata estimates 
```{r}
# URXBPH_LOG
I.genBmod1 <- svyglm(URXBPH_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXBP3_LOG
I.genBmod2 <- svyglm(URXBP3_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXTRS_LOG
I.genBmod3 <- svyglm(URXTRS_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXBUP_LOG
I.genBmod4 <- svyglm(URXBUP_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXEPB_LOG
I.genBmod5 <- svyglm(URXEPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXMPB_LOG
I.genBmod6 <- svyglm(URXMPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)

# URXPPB_LOG
I.genBmod7 <- svyglm(URXPPB_LOG ~ FOODSEC * SEX + URXUCR_final + AGEGROUP + EDUCAT + RACEETH + PIR_POLICY + cycleyr,
                     design = survey)
```


```{r}
emmeans::emm_options(rg.limit = 30000)

emms <- emmeans(I.genBmod1, ~ FOODSEC | SEX)
contrast(emms, method = "revpairwise")
```


```{r}

GENmodels_list<-list(I.genBmod1, I.genBmod2, I.genBmod3, I.genBmod4, I.genBmod5, I.genBmod6, I.genBmod7)
GENmod_names<-c("BPA", "BP3", "TRS", "BUP", "EPB", "MPB", "PPB")

emmeans::emm_options(rg.limit = 30000)

results <- lapply(seq_along(GENmodels_list), function(i) {
  model <- GENmodels_list[[i]]
  mod_name <- GENmod_names[i]
  
  emms <- emmeans(model, ~ FOODSEC | SEX)
  cont <- contrast(emms, method = "revpairwise")
  cont_summary <- summary(cont, infer = TRUE)
  
  # Exponentiate contrasts
  cont_summary$exp.estimate     <- exp(cont_summary$estimate)
  cont_summary$exp.lower.CL     <- exp(cont_summary$lower.CL)
  cont_summary$exp.upper.CL     <- exp(cont_summary$upper.CL)
  
  cont_summary$model <- mod_name
  cont_summary
})

GENall_results <- do.call(rbind, results)
write.csv(GENall_results, "Results/GENstrata_results2.csv", row.names=FALSE)
```

#PLOTS
###Model 3 adjusted signifcant findings
```{r}

RM_GGPlots <-read.csv("Results/RM_GGPlots2.csv")
names(RM_GGPlots) <- c("Model", "FOODSEC", "Food Security", "% Difference", "% Lower", "% Higher", "SE", "P_value")
brewer_pal <- brewer.pal(3, "Set1")
range(RM_GGPlots$`% Lower`, na.rm=TRUE)
range(RM_GGPlots$`% Higher`, na.rm=TRUE)



# Helper theme chunk to reuse in each plot
custom_theme <-
  theme(
    plot.title = element_text(hjust = 0.5, face="bold", size=18),
    panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),     
    plot.margin = margin(t = 10, r = 10, b = 25, l = 10)  # 
  )

# BPA
BPA_adj3 <- RM_GGPlots %>%
  filter(Model == "BPA", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
  ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.3, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BPA",
    x = "Food Security Levels",
    y = "% Difference in Concentrations") +
  labs(title = "BPA", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)  + 
  scale_y_continuous(limits = c(-42, 25))


# BP3
BP3_adj3 <- RM_GGPlots %>%
  filter(Model == "BP3", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
      ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "BP3", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+   scale_y_continuous(limits = c(-42, 25))


# TRS
TRS_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "TRS", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "TRS", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+   
  scale_y_continuous(limits = c(-42, 25))


# BUP
BUP_adj3 <- RM_GGPlots %>%
  filter(Model == "BUP", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
  ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "BUP", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
  scale_y_continuous(limits = c(-42, 25))


# EPB
EPB_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "EPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "EPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))


# MPB
MPB_adj3 <- ggplot(RM_GGPlots %>% filter(Model == "MPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)),
  aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)
) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "MPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))


#PPB
PPB_adj3 <- RM_GGPlots %>%
  filter(Model == "PPB", !is.na(`% Difference`), !is.na(`% Lower`), !is.na(`% Higher`)) %>%
 ggplot(aes(x = factor(`Food Security`, levels = c("High (Ref)", "Marginal", "Low", "Very Low")), y = `% Difference`)) +
  geom_point(size = 3, color = brewer_pal[1], alpha = 0.7) +
  geom_errorbar(aes(ymin = `% Lower`, ymax = `% Higher`), width = 0.2, color = brewer_pal[1]) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
  labs(title = "PPB", x = "Food Security Levels", y = "% Difference in Concentrations") +
  custom_theme +
  scale_x_discrete(drop = FALSE)+ 
    scale_y_continuous(limits = c(-42, 25))

```


```{r}

figure1<- ggarrange(
  BPA_adj3 + theme(legend.position="none"), 
  BP3_adj3 + theme(legend.position="none"), 
  TRS_adj3 + theme(legend.position="none"),
  BUP_adj3 + theme(legend.position="none"), 
  EPB_adj3 + theme(legend.position="none"), 
  MPB_adj3 + theme(legend.position="none"), 
  PPB_adj3 , # Leave legend ON only here
  ncol = 4, nrow = 2,
  labels = c("A", "B", "C", "D", "E", "G"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure1

```


## Stratified by gender plots

```{r}
GENstrata<-read.csv("Results/GENstrata2.csv")
brewer_pal <- brewer.pal(3, "Set1")

range(GENstrata$`X.Lower`, na.rm=TRUE) #65
range(GENstrata$`X.Higher`, na.rm=TRUE) #60



#create vector with race/ethnicity labels
sex_labels<- c("Female", "Male")
sex_colors <- c("Female" = brewer_pal[1], "Male" = brewer_pal[2])
names(sex_colors)<-sex_labels
sex_shapes <- c("Female" = 16, "Male" = 17) # Circle, triangle
names(sex_shapes)<-sex_labels



# MEP
BPA_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "BPA", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BPA",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BPA_adjEMM

# BP3
BP3_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "BP3", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BP3",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BP3_adjEMM

# TRS
TRS_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "TRS", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "TRS",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))  +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

TRS_adjEMM

# BUP
BUP_adjEMM<- ggplot(
  data = GENstrata %>%
    filter(model == "BUP", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "BUP",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

BUP_adjEMM

# EPB

EPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "EPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "EPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))  +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )
        


# MPB

MPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "MPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "MPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                   labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low"))  +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )

# PPB
PPB_adjEMM <- ggplot(
  data = GENstrata %>%
    filter(model == "PPB", 
           !is.na(X.Difference), 
           !is.na(X.Lower),
           !is.na(X.Higher)),
  aes(x = factor(FOODSEC, levels = c("1", "2", "3", "4")),
      y = X.Difference,
      color = SEX,
      shape = SEX)
) +
  geom_point(position = position_dodge(width = 0.4), size = 2.5, alpha = 0.7) +
  geom_errorbar(aes(ymin = X.Lower, ymax = X.Higher), 
                position = position_dodge(width = 0.4), width = 0.2) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +  
   labs(
    title = "PPB",
    x = "Food Security Levels",
    y = "% Difference in Concentrations", 
    color = "Sex", # title for color legend 
    shape = "Sex" # title for shape legend
  ) +
  scale_color_manual(values = sex_colors, breaks = sex_labels) +
  scale_shape_manual(values = sex_shapes, breaks = sex_labels) +
  scale_x_discrete(drop = FALSE, 
                  labels = c("1"= "High (Ref)", "2"="Marginal", "3"="Low", "4"="Very Low")) +
    scale_y_continuous(limits = c(-65, 60))+
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 18), 
     panel.grid.major.y = element_line(color = "grey90"),          
    panel.grid.minor = element_blank(),                          
    axis.line = element_line(color = "black"),                       
    legend.position = "bottom",                                      
    legend.title = element_text(size = 15, face = "bold"),           
    legend.text = element_text(size = 14),                           
    legend.box = "horizontal",                                      
    legend.key.size = unit(1.5, "lines"),                            
    strip.text = element_text(face = "bold", size = 16)             
  ) +
  guides(
    color = guide_legend(title.position = "left", nrow = 1, byrow = TRUE),
    shape = guide_legend(title.position = "left", nrow = 1, byrow = TRUE)
  )
PPB_adjEMM

ggexport(figure1, filename = "Results/figure1.pdf")

```


```{r}

# 
# BPA_adjEMM <- BPA_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62,
#            label = bquote(P[interaction] == .(formatC(0.08, digits=2, format="f"))),
#            size=5, hjust=0.5)
# BP3_adjEMM <- BP3_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62,
#            label = bquote(P[interaction] == .(formatC(0.08, digits=2, format="f"))),
#            size=5, hjust=0.5)
# 
# 
# TRS_adjEMM <- TRS_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62,
#           label = bquote(P[interaction] == .(formatC(0.02, digits=2, format="f"))),
#            size=5, hjust=0.5)
# 
# BUP_adjEMM <- BUP_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 63,
#            label = bquote(P[interaction] == .(formatC(0.0001, digits=4, format="f"))),
#            size=5, hjust=0.5)
# 
# 
# 
# EPB_adjEMM <- EPB_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62, 
#           label = bquote(P[interaction] == .(formatC(0.0002, digits=4, format="f"))),
#            size=5, hjust=0.5)
# 
# MPB_adjEMM <- MPB_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62, 
#          label = bquote(P[interaction] == .(formatC(0.30, digits=2, format="f"))),
#            size=5, hjust=0.5)
# 
# PPB_adjEMM <- PPB_adjEMM +
#   annotate("text", 
#            x = 2.5, y = 62, 
#            label = bquote(P[interaction] == .(formatC(0.01, digits=2, format="f"))),
#            size=5, hjust=0.5)

figure2<- ggarrange(
  BPA_adjEMM + theme(legend.position="none"),
  BP3_adjEMM + theme(legend.position="none"), 
  TRS_adjEMM + theme(legend.position="none"), 
  BUP_adjEMM + theme(legend.position="none"), 
  EPB_adjEMM + theme(legend.position="none"), 
  MPB_adjEMM + theme(legend.position="none"), 
  PPB_adjEMM,  # Leave legend ON only here
  ncol = 4, nrow = 2,
  labels = c("A", "B", "C", "D", "E", "G"),
  common.legend = TRUE, # Should propagate the legend from one plot
  legend = "bottom"
)

figure2

ggexport(figure2, filename = "Results/figure2.pdf")
```







