---
title: "Data_Recoding"
author: "Ayishat Yussuf"
date: "2025-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```
# Load Packages and Datasets

```{r}
library(readr) #read in csv files 
library(survey) #analysis for survey data (e.g NHANES)
library(magrittr) #pip operations (%>%) 
library(dplyr) #data manipulation (filter, select, mutate etc) 
library(ggplot2) #creating data visualizations/plots
```

#Load File
```{r}
FullData_allcycles<-read.csv("~/Documents/Github/Phenols-Analysis/Phenols_DataUnclean/FullData_allcycles.csv")

names(FullData_allcycles)
```
Note to Ayishat: Initial Population: 16,349
##Food Security
```{r}
#fsdhh
table(FullData_allcycles$FSDHH, useNA="ifany")
#198 missing

#Variables
FullData_allcycles<-FullData_allcycles%>%
  mutate(FOODSEC=case_when(
    FSDHH==1 ~1,  #fully food secure (ref)
    FSDHH==2 ~2, #marginally food secure (1-2)
    FSDHH==3 ~3, # low food security 
    FSDHH==4 ~4,  #very low food security 
    TRUE~NA_real_
  ))

table(FullData_allcycles$FOODSEC,useNA="ifany") #198 missing
```
Note to Ayishat: 198 missing food security 
##Age
```{r}
summary(FullData_allcycles$RIDAGEYR, useNA="ifany") #no missing

class(FullData_allcycles$RIDAGEYR) #to see what type of variable you have 

FullData_allcycles<-FullData_allcycles%>%
  filter(FullData_allcycles$RIDAGEYR>=18) #filtered for observations Age greater than or equal to 18

quartiles<-quantile(FullData_allcycles$RIDAGEYR)
print(quartiles)

FullData_allcycles$AGEGROUP<-cut(
  FullData_allcycles$RIDAGEYR,
  breaks=quartiles, 
  include.lowest=TRUE,
  labels= c("18-31", "32-46", "47-63", "64-85")
)

#no missing
table(FullData_allcycles$AGEGROUP, useNA="ifany")
```
decided to keep male as reference - september 19th, 2025

Note to Ayishat: Following removal of children, sample size is now 11,479 observations

##Sex
```{r}
table(FullData_allcycles$RIAGENDR, useNA="ifany")

FullData_allcycles<-FullData_allcycles%>%
  mutate(SEX=ifelse(RIAGENDR==1, 1, 2))

levels(FullData_allcycles$SEX)<-(c("1", "2"))

table(FullData_allcycles$SEX, useNA="ifany") #no NA
```


##Race/Eth
```{r}
#Recode race/ethnicity 
table(FullData_allcycles$RIDRETH1, useNA="ifany")

FullData_allcycles<-FullData_allcycles%>%
  mutate(RACEETH=case_when(
    RIDRETH1==3 ~1,
    RIDRETH1==4 ~2, 
    RIDRETH1==1 ~3, 
    RIDRETH1==2 ~4, 
    RIDRETH1==5 ~5, 
    TRUE~NA_real_
  ))


table(FullData_allcycles$RACEETH, useNA="ifany") #no NA
```


##Education
```{r}
#education 
summary(FullData_allcycles$DMDEDUC3)
summary(FullData_allcycles$DMDEDUC3)
FullData_allcycles <- FullData_allcycles %>%
  mutate(EDUCAT = case_when(
   DMDEDUC3< 13 & RIDAGEYR<20 ~ 1,
    DMDEDUC2 < 3 ~ 1, 
    DMDEDUC2== 1 ~2, 
    DMDEDUC3 %in% c(13, 14, 15) ~ 2, 
    DMDEDUC2 == 3 ~ 2,
    DMDEDUC2 == 4 ~ 3,
    DMDEDUC2 == 5 ~ 4,
    TRUE ~ NA_real_
  ))

table(FullData_allcycles$EDUCAT, useNA="ifany") #23 NA
```

##Poverty
```{r}
#poverty
summary(FullData_allcycles$INDFMPIR, useNA="ifany") #1030 missing

FullData_allcycles<-FullData_allcycles%>%
  mutate(PIRCAT=case_when(
    INDFMPIR<=0.99 ~1, 
    INDFMPIR>=1 & INDFMPIR<2 ~2, 
    INDFMPIR>=2 & INDFMPIR<3 ~3, 
    INDFMPIR>=3 & INDFMPIR<5 ~4, 
    INDFMPIR>=5 ~5, 
    is.na(INDFMPIR)~NA_real_
  ))

table(FullData_allcycles$PIRCAT, useNA="ifany") #1030 
```
1030 missing PIR

##CycleYr
Check levels of cycle year, make sure they match up with your analysis time span, and check missingness
```{r}
#cycleyears
table(FullData_allcycles$cycleyr, useNA="ifany")

```


##Phenols
Urinary Bisphenol A - URXBPH
Urinary Bisphenol F - URXBPF
Urinary Bisphenol S - URXBPS
Urinary 2-Hydroxy-4-metoxybenzophenone (Benzophenone-3) - URXBP3
Urinary 2,4,4’-Trichloro-2’-hydroxyphenyl ether (Triclosan) - URXTRS
Urinary Butyl paraben - URXBUP
Urinary Propyl paraben - URXPPB
Urinary Ethyl paraben - URXEPB
Urinary Methyl paraben - URXMPB
```{r}
names(FullData_allcycles)

sum(is.na(FullData_allcycles$URXUCR.y))#9685


FullData_allcycles$URXUCR_final <- coalesce(
  as.numeric(FullData_allcycles$URXUCR),
  as.numeric(FullData_allcycles$URXUCR.x),
  as.numeric(FullData_allcycles$URXUCR.y)
)

summary(FullData_allcycles$URXUCR_final, useNA="ifany") #300 missing creatinine 

```

###Create Sample and Sample 2 (final analytic dataset)

Marshae Note to Ayishat: Before we generate sample: prior to any missingness removal we have 13207 participants
```{r}

#generate sample (samplepopulation) from nonmissing in exposure/covariates
FullData_allcycles<-FullData_allcycles %>%
  mutate(SAMPLE= ifelse(is.na(EDUCAT)| is.na(PIRCAT) | is.na(SEX)| is.na(RACEETH) | is.na(AGEGROUP)| is.na(FOODSEC),0,1))

sum(is.na(FullData_allcycles$EDUCAT) | 
    is.na(FullData_allcycles$PIRCAT) | 
    is.na(FullData_allcycles$SEX) | 
    is.na(FullData_allcycles$RACEETH) | 
    is.na(FullData_allcycles$AGEGROUP) | 
    is.na(FullData_allcycles$FOODSEC) | 
    is.na(FullData_allcycles$cycleyr)) #1056 missing 

table(FullData_allcycles$SAMPLE, useNA="ifany") #10,423

#second sampling variable (missing in outcome, creatinine)
FullData_allcycles$SAMPLE2<- FullData_allcycles$SAMPLE

names(FullData_allcycles)#check if variable was created

saveRDS(FullData_allcycles, "FullData_allcycles.rds") #run this to save dataset with changes/recoding 


```
Note to Ayishat: Following removal we have 10,423 with exclusion of missing covariates, and exposure data





###Outliers

raw column= phenol concentration (ng/mL), flag_column= below or above LOD?  LOD= obtained from NHANES documentation (the doc file on NHANES website. You can use last cycle years doc to fill in these values)

Urinary Bisphenol A - URXBPH
Urinary Bisphenol F - URXBPF--EXCLUDE (10/8/25)
Urinary Bisphenol S - URXBPS --EXCLUDE (10/8/25)
Urinary 2-Hydroxy-4-metoxybenzophenone (Benzophenone-3) - URXBP3
Urinary 2,4,4’-Trichloro-2’-hydroxyphenyl ether (Triclosan) - URXTRS
Urinary Butyl paraben - URXBUP
Urinary Propyl paraben - URXPPB
Urinary Ethyl paraben - URXEPB
Urinary Methyl paraben - URXMPB

Missingness schema for outcome/creatinine, 
Remove those who have missing data
Remove those who are outliers (5 standard deviations above the mean)


*MARSHAE NOTES TO AYISHAT: 
--2003-2004: BPA, Triclsoan, BP3
--2005-2006: Phenols + Parabens (introduced this year): BPA, BP3, Triclosan, Methyl, Ethyl, Propyl, Butyl Paraben 

Decision: Too restrictive in list previous list. 
Will start from 2005-2005 and only include: 
  BPA, BP3, Triclosan, Butyl, Ethyl, Methyl, Propyl Parabens
  
  
```{r}
FullData_allcycles<-readRDS("FullData_allcycles.rds")
```


###Summary

```{r}
table(FullData_allcycles$URDBPHLC)#870 below LOD
table(FullData_allcycles$URDBP3LC)#430 below LOD
table(FullData_allcycles$URDTRSLC)#2816 below LOD
table(FullData_allcycles$URDBUPLC)#7032 below LOD
table(FullData_allcycles$URDEPBLC)#5590 below LOD
table(FullData_allcycles$URDMPBLC)#88 below LOD
table(FullData_allcycles$URDPPBLC)#536 below LOD

```

Note to Ayishat: quite a bit of values below LOD, namely ethyl and butyl paraben and TRS
confirmation that values below LLOD have already been replaced by NHANES

```{r}
BPH<- FullData_allcycles[FullData_allcycles$URDBPHLC == 1, ]
BP3<- FullData_allcycles[FullData_allcycles$URDBP3LC == 1, ]
TRS<- FullData_allcycles[FullData_allcycles$URDTRSLC == 1, ]
BUP<- FullData_allcycles[FullData_allcycles$URDBUPLC == 1, ]
EPB<- FullData_allcycles[FullData_allcycles$URDEPBLC == 1, ]
MPB<- FullData_allcycles[FullData_allcycles$URDMPBLC == 1, ]
PPB<- FullData_allcycles[FullData_allcycles$URDPPBLC == 1, ]

summary(BPH$URXBPH)
summary(BP3$URXBP3)
summary(TRS$URXTRS)
summary(BUP$URXBUP)
summary(EPB$URXEPB)
summary(MPB$URXMPB)
summary(PPB$URXPPB)
```


```{r}
#BPH
summary(FullData_allcycles$URXBPH, na.rm = FALSE) #387 missing
#BP3
summary(FullData_allcycles$URXBP3, na.rm = FALSE) #387 missing
#TRS
summary(FullData_allcycles$URXTRS, na.rm = FALSE) #387 missing
#BUP
summary(FullData_allcycles$URXBUP, na.rm = FALSE) #387 missing
#EPB
summary(FullData_allcycles$URXEPB, na.rm = FALSE) #387 missing
#MPB
summary(FullData_allcycles$URXMPB, na.rm = FALSE) #387 missing
#PPB
summary(FullData_allcycles$URXPPB, na.rm = FALSE) #387 missing
```
Note to Ayishat: low missingness in Phenols

```{r}
# FullData_allcycles <- FullData_allcycles %>%
#   mutate(SAMPLE2 = ifelse(SAMPLE == 1 &
#                           URDBPHLC != 1 &
#                           URDBP3LC != 1 &
#                           URDTRSLC != 1 &
#                           URDBUPLC != 1 &
#                           URDEPBLC != 1 &
#                           URDMPBLC != 1 &
#                           URDPPBLC != 1, 1, 0))
# 
# table(FullData_allcycles$SAMPLE2) #2139 final sample size 


FullData_allcycles2 <- FullData_allcycles %>%
  mutate(SAMPLE2 = ifelse(SAMPLE == 1 &
                          (URDBPHLC != 1 |
                           URDBP3LC != 1 |
                           URDTRSLC != 1 |
                           URDBUPLC != 1 |
                           URDEPBLC != 1 |
                           URDMPBLC != 1 |
                           URDPPBLC != 1) & 
                            !is.na(URXUCR_final), 1, 0))


table(FullData_allcycles2$SAMPLE2) #10094 final sample size, only 3 excluded missing creatinine

idx <- !is.na(FullData_allcycles2$URXUCR_final)
table(FullData_allcycles2$SAMPLE2[idx] == 1, useNA = "ifany") #75 missing ?

# After you've created SAMPLE2
FullData_allcycles2$SAMPLE2[is.na(FullData_allcycles2$SAMPLE2)] <- 0

table(FullData_allcycles2$SAMPLE2, useNA = "ifany") #10094


```


note to ayishat: 2139 due to large number of values below LLOD in parabens

###Visualizations
####NonLog
```{r}
boxplot(FullData_allcycles$URXBPH[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXBP3[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXTRS[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXBUP[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXEPB[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXMPB[FullData_allcycles$SAMPLE2 == 1])
boxplot(FullData_allcycles$URXPPB[FullData_allcycles$SAMPLE2 == 1])

```
####Log
```{r}

boxplot(log(FullData_allcycles$URXBPH[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXBP3[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXTRS[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXBUP[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXEPB[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXMPB[FullData_allcycles$SAMPLE2 == 1]))
boxplot(log(FullData_allcycles$URXPPB[FullData_allcycles$SAMPLE2 == 1]))
```




```{r}
# Example: Suppose you join on 'cycleyr'
# FullData_allcycles$cycleyr <- as.character(FullData_allcycles$cycleyr)
# lod_long$cycleyr <- as.character(lod_long$cycleyr)
# 
# for (metab in names(metabolites)) {
#   raw_col <- metabolites[[metab]]$raw_col
#   flag_col <- metabolites[[metab]]$flag_col
# 
#   FullData_allcycles[[metab]] <- FullData_allcycles[[raw_col]]
# 
#   lod_vector <- dplyr::left_join(
#     data.frame(cycleyr = FullData_allcycles$cycleyr),
#     lod_long[lod_long$metab == metab, c("cycleyr", "lod")],
#     by = "cycleyr"
#   )$lod
# 
#   mask <- !is.na(FullData_allcycles[[flag_col]]) & FullData_allcycles[[flag_col]] == 1
#   FullData_allcycles[[metab]][mask] <- lod_vector[mask] / sqrt(2)
# 
#   metab_sd <- sd(FullData_allcycles[[metab]][FullData_allcycles$SAMPLE == 1], na.rm = TRUE)
#   cutoff <- 5 * metab_sd
#   FullData_allcycles$SAMPLE2[FullData_allcycles[[metab]] > cutoff] <- 0
# 
#   cat("\n---", metab, "---\n")
#   cat("Standard Deviation:", round(metab_sd, 3), "\n")
#   cat("5x Standard Deviation Cutoff:", round(cutoff, 3), "\n")
#   cat("Number of Outliers (SAMPLE2 = 0):", sum(FullData_allcycles$SAMPLE2 == 0, na.rm = TRUE), "\n")
#   cat("Post-Outlier Standard Deviation:", 
#       round(sd(FullData_allcycles[[metab]][FullData_allcycles$SAMPLE2 == 1], na.rm = TRUE), 3), "\n")
# }
```




*Marshae Note to self: RESOLVED
The detection limits were constant for all of the analytes in the data set. Two variables are provided for each of these analytes. The variable name ending in “LC” (ex., URDBPHLC) indicates whether the result was below the limit of detection: the value “0” means that the result was at or above the limit of detection, “1” indicates that the result was below the limit of detection. For analytes with analytic results below the lower limit of detection (ex., URDBPHLC=1), an imputed fill value was placed in the analyte results field. This value is the lower limit of detection divided by square root of 2 (LLOD/sqrt[2]). The other variable prefixed URX (ex., URXBPH) provides the analytic result for that analyte.---2015/2016 documentation NHANES





##Weights

####fulldataset
```{r}
#change survey weights 
# FullData_allcycles <- FullData_allcycles %>%
#   mutate(SURVWEIGHT = case_when(
#     cycleyr >= 1 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
#     cycleyr >= 2 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
#     cycleyr >= 3 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
#     cycleyr >= 4 & !is.na(WTSA2YR) ~ WTSA2YR * (1 / 6),
#     cycleyr >= 5 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
#     cycleyr >= 6 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
#     TRUE ~ NA_real_
#   ))
# 
# 
# write_csv(FullData_allcycles, "~/Documents/Github/Phenols-Analysis/Phenols_DataClean/FullData_allcyclesFINAL.csv")
# 
# saveRDS(FullData_allcycles, "~/Documents/Github/Phenols-Analysis/Phenols_DataClean/FullData_allcyclesFINAL.rds")
```





####E/Bexclusion
```{r}
#change survey weights 
FullData_allcycles2 <- FullData_allcycles2 %>%
  mutate(SURVWEIGHT = case_when(
    cycleyr >= 1 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
    cycleyr >= 2 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
    cycleyr >= 3 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
    cycleyr >= 4 & !is.na(WTSA2YR) ~ WTSA2YR * (1 / 6),
    cycleyr >= 5 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
    cycleyr >= 6 & !is.na(WTSB2YR) ~ WTSB2YR * (1 / 6),
    TRUE ~ NA_real_
  ))


with(FullData_allcycles2,
     table(SAMPLE2 == 1, is.na(SURVWEIGHT), useNA = "ifany")) #no one in sample has missing 

write_csv(FullData_allcycles2, "~/Documents/Github/Phenols-Analysis/Phenols_DataClean/FullData_allcyclesFINAL2.csv")

saveRDS(FullData_allcycles2, "~/Documents/Github/Phenols-Analysis/Phenols_DataClean/FullData_allcyclesFINAL2.rds")
```




